---
title: "Viz Tutorial"
output: html_notebook
---

#Tutorial

```{r}
library(tidyverse)
library(ggplot2)
library(readr)
```


## Read Data

```{r}

#In English, Spanish, French, German all profession words occur in the corpus
df_english <- read_csv("english_results.csv")
#df2 <- read_csv("arabic_results2.csv")
df_spanish <- read_csv("spanish_even.csv")
df_german <- read_csv("german_even.csv")
df_french <- read_csv("french_even.csv")

#TODO-I'm not sure I am choosing the right Chinese results column
df_chinese <- read_csv("chinese_results_pca0.csv")
df_chinese_pca1 <- read_csv("chinese_results_pca1.csv")

#in Farsi journalist, astronomer and biologist do not occur in corpus
df_farsi <- read_csv("farsi_results.csv")

#in Arabic, filmmaker, astronaut and astronomer do not occur in the corpus
df_arabic <- read_csv("arabic_even.csv")

#in Urdu comedian, adventurer, analyst, astronomer and biologist do not occur in the corpus
#Also we have English words transcribed in Arabic character set for 9 words - could those be considered neutral words?
#Only 2 female profession variants occur in the corpus - teacher and actor
df_urdu <- read_csv("urdu_even.csv")

#In Wolof only two profession words - person and musician appear in the corpus
df_wolof <- read_csv("wolof_results.csv")


#For Gendered languages, read in female, male and neutral separately
df_spanish_female <-read_csv("spanish_female.csv")
df_spanish_male <-read_csv("spanish_male.csv")
df_spanish_neutral <-read_csv("spanish_neutral.csv")
df_spanish_weighted <-read_csv("spanish_weighted.csv")
df_spanish_weighted_simple <-read_csv("spanish_weighted.csv")

df_french_female <-read_csv("french_female.csv")
df_french_male <-read_csv("french_male.csv")
df_french_neutral <-read_csv("french_neutral.csv")
df_french_weighted <-read_csv("french_weighted.csv")
df_french_weighted_simple <-read_csv("french_weighted.csv")

df_german_female <-read_csv("german_female.csv")
df_german_male <-read_csv("german_male.csv")
df_german_neutral <-read_csv("german_neutral.csv")
df_german_weighted <-read_csv("german_weighted.csv")
df_german_weighted_simple <-read_csv("german_weighted.csv")

#Note Arabic character's lost in CSV
df_arabic_female <-read_csv("arabic_female.csv")
df_arabic_male <-read_csv("arabic_male.csv")
df_arabic_neutral <-read_csv("arabic_neutral.csv")
df_arabic_weighted <-read_csv("arabic_weighted.csv")
df_arabic_weighted_simple <-read_csv("arabic_weighted.csv")


df_urdu_female <-read_csv("urdu_female.csv")
df_urdu_male <-read_csv("urdu_male.csv")
df_urdu_neutral <-read_csv("urdu_neutral.csv")
df_urdu_english <-read_csv("urdu_english.csv")
df_urdu_weighted <-read_csv("urdu_weighted.csv")
df_urdu_weighted_simple <-read_csv("urdu_weighted.csv")
```


## Label data
```{r}
#Add language variable in each 
df_english <- df_english %>% 
  mutate(language = "English")

df_farsi <- df_farsi %>%
  mutate(language = "Farsi")

df_chinese <- df_chinese %>% 
  mutate(language = "Chinese")


df_chinese_pca0 <- df_chinese %>% 
    mutate(gender = "PCA0")
df_chinese_pca1 <- df_chinese_pca1 %>% 
  mutate(language = "Chinese")%>%
  mutate(gender = "PCA1")


df_wolof <- df_wolof %>%
  mutate(language = "Wolof")


df_arabic <- df_arabic %>% 
  mutate(language = "Arabic")
df_arabic_combo <- df_arabic %>%
  mutate(Arabic = "EMPTY") %>%
  mutate(gender = "Weighted Equally")
df_arabic_weighted_simple <- df_arabic_weighted_simple %>%
   mutate(language = "Arabic")
df_arabic_weighted <- df_arabic_weighted %>%
  mutate(language = "Arabic")%>%
  mutate(Arabic = "EMPTY") %>%
  mutate(gender = "Weighted By Word Count")
df_arabic_female <- df_arabic_female %>%
  mutate(language = "Arabic") %>%
  mutate(gender = "Female")
df_arabic_male <- df_arabic_male %>%
  mutate(language = "Arabic")%>%
  mutate(gender = "Male")
df_arabic_neutral <- df_arabic_neutral %>%
  mutate(language = "Arabic")%>%
  mutate(gender = "Neutral")



df_spanish <- df_spanish %>%
  mutate(language = "Spanish")
df_spanish_combo <- df_spanish %>%
  mutate(Spanish = "EMPTY") %>%
  mutate(gender = "Weighted Equally")
df_spanish_weighted_simple <- df_spanish_weighted_simple %>%
   mutate(language = "Spanish")
df_spanish_weighted <- df_spanish_weighted %>%
  mutate(language = "Spanish")%>%
  mutate(Spanish = "EMPTY") %>%
  mutate(gender = "Weighted By Word Count")
df_spanish_female <- df_spanish_female %>%
  mutate(language = "Spanish") %>%
  mutate(gender = "Female")
df_spanish_male <- df_spanish_male %>%
  mutate(language = "Spanish")%>%
  mutate(gender = "Male")
df_spanish_neutral <- df_spanish_neutral %>%
  mutate(language = "Spanish")%>%
  mutate(gender = "Neutral")


df_french <- df_french %>%
  mutate(language = "French")
df_french_combo <- df_french %>%
  mutate(French = "EMPTY") %>%
  mutate(gender = "Weighted Equally")
df_french_weighted_simple <- df_french_weighted_simple %>%
   mutate(language = "French")
df_french_weighted <- df_french_weighted %>%
  mutate(language = "French")%>%
  mutate(French = "EMPTY") %>%
  mutate(gender = "Weighted By Word Count")
df_french_female  <- df_french_female  %>%
  mutate(language = "French")%>%
  mutate(gender = "Female")
df_french_male  <- df_french_male %>%
  mutate(language = "French")%>%
  mutate(gender = "Male")
df_french_neutral <- df_french_neutral %>%
  mutate(language = "French")%>%
  mutate(gender = "Neutral")

df_german <- df_german %>%
  mutate(language = "German")
df_german_combo <- df_german %>%
  mutate(German = "EMPTY") %>%
  mutate(gender = "Weighted Equally")
df_german_weighted_simple <- df_german_weighted_simple %>%
   mutate(language = "German")
df_german_weighted <- df_german_weighted %>%
  mutate(language = "German")%>%
  mutate(German = "EMPTY") %>%
  mutate(gender = "Weighted By Word Count")
df_german_female <- df_german_female %>%
  mutate(language = "German")%>%
  mutate(gender = "Female")
df_german_male <- df_german_male %>%
  mutate(language = "German")%>%
  mutate(gender = "Male")
df_german_neutral <- df_german_neutral %>%
  mutate(language = "German")%>%
  mutate(gender = "Neutral")


df_urdu <- df_urdu %>% 
  mutate(language = "Urdu")
df_urdu_combo <- df_urdu %>%
  mutate(Urdu = "EMPTY") %>%
  mutate(gender = "Weighted Equally")
df_urdu_weighted_simple <- df_urdu_weighted_simple %>%
   mutate(language = "Urdu")
df_urdu_weighted <- df_urdu_weighted %>%
  mutate(language = "Urdu")%>%
  mutate(Urdu = "EMPTY") %>%
  mutate(gender = "Weighted By Word Count")
df_urdu_female <- df_urdu_female %>%
  mutate(language = "Urdu")%>%
  mutate(gender = "Female")
df_urdu_male <- df_urdu_male %>%
  mutate(language = "Urdu") %>%
  mutate(gender = "Male")
df_urdu_neutral <- df_urdu_neutral %>%
  mutate(language = "Urdu")%>%
  mutate(gender = "Urdu Neutral")
df_urdu_english <- df_urdu_english %>%
  mutate(language = "Urdu")%>%
  mutate(gender = "English Neutral")
```

#Combine data
```{r}
df_total <-  rbind(df_english, df_chinese, df_arabic, df_spanish, df_german, df_farsi, df_french, df_urdu, df_wolof)

df_total_weighted <-  rbind(df_english, df_chinese, df_arabic_weighted_simple, df_farsi, df_french_weighted_simple, df_urdu_weighted_simple, df_wolof, df_spanish_weighted_simple, df_german_weighted_simple)

df_gendered_languages <-rbind(df_spanish, df_german, df_arabic, df_french, df_urdu)

df_gendered_languages_weighted <-rbind(df_spanish_weighted_simple, df_german_weighted_simple, df_french_weighted_simple, df_arabic_weighted_simple, df_urdu_weighted_simple)

df_nongendered_languages<-rbind(df_english, df_chinese, df_farsi, df_wolof)

df_spanish_all <-rbind(df_spanish_female, df_spanish_male, df_spanish_neutral, 
                       df_spanish_combo, df_spanish_weighted)
df_french_all <-rbind(df_french_female, df_french_male, df_french_neutral, 
                      df_french_combo, df_french_weighted)
df_german_all <-rbind(df_german_female, df_german_male, df_german_neutral,
                      df_german_combo, df_german_weighted)

df_arabic_all <-rbind(df_arabic_female, df_arabic_male, df_arabic_neutral,
                      df_arabic_combo, df_arabic_weighted)
df_urdu_all <-rbind(df_urdu_female, df_urdu_male, df_urdu_neutral, df_urdu_english, df_urdu_combo, df_urdu_weighted)

df_chinese_all <-rbind(df_chinese_pca0, df_chinese_pca1)
```

#Visualizations


### Basic bar chart
```{r}
ggplot(data = df_english, aes(x = professions, y = gender_bias)) +
  geom_bar(stat = "identity") 
```

### Add color for bias 

```{r}
ggplot(data = df_english, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity")
```

###Flip coordinates

```{r}
ggplot(data = df_english, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip()
```

### Add title/axis label
```{r}
ggplot(data = df_english, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Set (English Only)",
       y = "Gender Bias",
       x = element_blank())
```


### Change color
```{r}
ggplot(data = df_english, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Set (English Only)",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "dodgerblue2", "firebrick2"))
```

### Change axis to be balanced
```{r}
ggplot(data = df_english, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Set (English Only)",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "dodgerblue2", "firebrick2")) +
  expand_limits(y = c(-.3, 0.3))
```


### Mess with "Theme" (delete legend, change font, center title) - FINAL VERSION in ENGLISH?
```{r}
pEnglish=ggplot(data = df_english, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Set (English)",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "dodgerblue2", "firebrick2")) +
  expand_limits(y = c(-.3, 0.3)) + 
  theme(
  legend.position = "none",
  text = element_text(family = "Gill Sans"),
  plot.title = element_text(hjust = 0.5)
)
ggsave(file = "Professions_English.png", pEnglish)
```

### Use same settings in Spanish (Weighted By Word Count)
```{r}
ggplot(data = df_spanish_weighted_simple, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Set (Spanish)",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "dodgerblue2", "firebrick2")) +
  expand_limits(y = c(-.3, 0.3)) + 
  theme(
  legend.position = "none",
  text = element_text(family = "Gill Sans"),
  plot.title = element_text(hjust = 0.5)
)
```
### Use same settings in French (Weighted By Word Count)
```{r}
ggplot(data = df_french_weighted_simple, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Set (French)",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "dodgerblue2", "firebrick2")) +
  expand_limits(y = c(-.3, 0.3)) + 
  theme(
  legend.position = "none",
  text = element_text(family = "Gill Sans"),
  plot.title = element_text(hjust = 0.5)
)
```
### Use same settings in German (Weighted By Word Count)
```{r}
ggplot(data = df_german_weighted_simple, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Set (German)",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "dodgerblue2", "firebrick2")) +
  expand_limits(y = c(-.3, 0.3)) + 
  theme(
  legend.position = "none",
  text = element_text(family = "Gill Sans"),
  plot.title = element_text(hjust = 0.5)
)
```
## Use same settings in Arabic (Weighted By Word Count)
```{r}
ggplot(data = df_arabic_weighted_simple, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Set (Arabic)",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "dodgerblue2", "firebrick2")) +
  expand_limits(y = c(-.3, 0.3)) + 
  theme(
  legend.position = "none",
  text = element_text(family = "Gill Sans"),
  plot.title = element_text(hjust = 0.5)
)
```
## Use same settings in Urdu (Weighted By Word Count)
```{r}
ggplot(data = df_urdu_weighted_simple, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Set (Urdu)",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "dodgerblue2", "firebrick2")) +
  expand_limits(y = c(-.3, 0.3)) + 
  theme(
  legend.position = "none",
  text = element_text(family = "Gill Sans"),
  plot.title = element_text(hjust = 0.5)
)
```
## Use same settings in Chinese (PCA0)
```{r}
pChinese_PCA0=ggplot(data = df_chinese, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Set (Chinese)",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "dodgerblue2", "firebrick2")) +
  expand_limits(y = c(-.3, 0.3)) + 
  theme(
  legend.position = "none",
  text = element_text(family = "Gill Sans"),
  plot.title = element_text(hjust = 0.5)
)
ggsave(file = "Professions_Chinese0.png", pChinese_PCA0)
```


## Use same settings in Farsi
```{r}
pFarsi=ggplot(data = df_farsi, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Set (Farsi)",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "dodgerblue2", "firebrick2")) +
  expand_limits(y = c(-.3, 0.3)) + 
  theme(
  legend.position = "none",
  text = element_text(family = "Gill Sans"),
  plot.title = element_text(hjust = 0.5)
)
ggsave(file = "Professions_Farsi.png", pFarsi)
```

## Use same settings in Wolof
```{r}
pWolof= ggplot(data = df_wolof, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Set (Wolof)",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "firebrick2", "dodgerblue2")) +
  expand_limits(y = c(-.3, 0.3)) + 
  theme(
  legend.position = "none",
  text = element_text(family = "Gill Sans"),
  plot.title = element_text(hjust = 0.5)
)
ggsave(file = "Professions_Wolof.png", pWolof)
```
### All languages

```{r}
pAllEven=ggplot(data = df_total, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Sets Across All Languages (Weighted Evenly)",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "dodgerblue2", "firebrick2")) +
  expand_limits(y = c(-.3, 0.3)) + 
  theme(
  legend.position = "none",
  text = element_text(family = "Gill Sans"),
  plot.title = element_text(hjust = 0.5),
  axis.text.y=element_blank(),
        axis.ticks.y=element_blank()
) +
 facet_wrap(~language, nrow = 3)
pAllEven
ggsave(file = "Professions_AllEven.png", pAllEven)
```

### All languages (Weighted average)

```{r}
pAllWeighted=ggplot(data = df_total_weighted, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Sets Across All Languages (Weighted By Word Count)",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "dodgerblue2", "firebrick2")) +
  expand_limits(y = c(-.3, 0.3)) + 
  theme(
  legend.position = "none",
  text = element_text(family = "Gill Sans"),
  plot.title = element_text(hjust = 0.5),
  axis.text.y=element_blank(),
        axis.ticks.y=element_blank()
) +
 facet_wrap(~language, nrow = 3)
ggsave(file = "Professions_AllWeighted.png", pAllWeighted)
```

# Put English, Chinese, Farsi, Wolof together
```{r}
pNonGendered= ggplot(data = df_nongendered_languages, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Languages Without Grammatically Gendered Nouns",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "dodgerblue2", "firebrick2")) +
  expand_limits(y = c(-.3, 0.3)) + 
  theme(
  legend.position = "none",
  text = element_text(family = "Gill Sans"),
  plot.title = element_text(hjust = 0.5)
) +
 facet_wrap(~language, nrow = 2)
ggsave(file = "Professions_NonGendered.png", pNonGendered)
```

# Put French, German, Spanish, Arabic together

```{r}
ggplot(data = df_gendered_languages, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Sets (Gendered Languages - Weighted Equally)",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "dodgerblue2", "firebrick2")) +
  expand_limits(y = c(-.3, 0.3)) + 
  theme(
  legend.position = "none",
  text = element_text(family = "Gill Sans"),
  plot.title = element_text(hjust = 0.5)
) +
 facet_wrap(~language, nrow = 2)
```
# Put French, German, Spanish, Arabic together

```{r}
ggplot(data = df_gendered_languages_weighted, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Set (Gendered Languages - Weighted Average)",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "dodgerblue2", "firebrick2")) +
  expand_limits(y = c(-.3, 0.3)) + 
  theme(
  legend.position = "none",
  text = element_text(family = "Gill Sans"),
  plot.title = element_text(hjust = 0.5)
) +
 facet_wrap(~language, nrow = 2)
```

# Graph of just the ones for gendered languages show each variant (just female, just male, just neutral) - SPANISH
#NA lines come from empty professions
#why are they sorted like they are?  how about alphabetical vs reverse alphabetical?
#When I tried to use Spanish words, Things with accents are missing :-(
```{r}
pSpanish = ggplot(data = df_spanish_all, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Set (Spanish)",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "dodgerblue2", "firebrick2")) +
  expand_limits(y = c(-.3, 0.3)) + 
  theme(
  legend.position = "none",
  text = element_text(family = "Gill Sans"),
  plot.title = element_text(hjust = 0.5)
) +
 facet_wrap(~gender, nrow = 2)
ggsave(file = "Professions_Spanish.png", pSpanish)
```





# Graph of just the ones for gendered languages show each variant (just female, just male, just neutral) - FRENCH

```{r}
pFrench= ggplot(data = df_french_all, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Set (French)",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "dodgerblue2", "firebrick2")) +
  expand_limits(y = c(-.3, 0.3)) + 
  theme(
  legend.position = "none",
  text = element_text(family = "Gill Sans"),
  plot.title = element_text(hjust = 0.5)
) +
 facet_wrap(~gender, nrow = 2)
ggsave(file = "Professions_French.png", pFrench)
```


# Graph of just the ones for gendered languages show each variant (just female, just male, just neutral) - GERMAN
```{r}
pGerman= ggplot(data = df_german_all, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Set (German)",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "dodgerblue2", "firebrick2")) +
  expand_limits(y = c(-.3, 0.3)) + 
  theme(
  legend.position = "none",
  text = element_text(family = "Gill Sans"),
  plot.title = element_text(hjust = 0.5)
) +
 facet_wrap(~gender, nrow = 2)
ggsave(file = "Professions_German.png", pGerman)
```


# Graph of just the ones for gendered languages show each variant (just female, just male, just neutral) - ARABIC
```{r}
pArabic= ggplot(data = df_arabic_all, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Set (Arabic)",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "dodgerblue2", "firebrick2")) +
  expand_limits(y = c(-.3, 0.3)) + 
  theme(
  legend.position = "none",
  text = element_text(family = "Gill Sans"),
  plot.title = element_text(hjust = 0.5)
) +
 facet_wrap(~gender, nrow = 2)
ggsave(file = "Professions_Arabic.png", pArabic)
```
# Graph of just the ones for gendered languages show each variant (just female, just male, just neutral) - URDU
```{r}
pUrdu= ggplot(data = df_urdu_all, aes(x = professions, y = gender_bias, fill = pos)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Profession Set (Urdu)",
       y = "Gender Bias",
       x = element_blank()) +
  scale_fill_manual(values = c( "dodgerblue2", "firebrick2")) +
  expand_limits(y = c(-.3, 0.3)) + 
  theme(
  legend.position = "none",
  text = element_text(family = "Gill Sans"),
  plot.title = element_text(hjust = 0.5)
) +
 facet_wrap(~gender, nrow = 2)
ggsave(file = "Professions_Urdu.png", pUrdu)
```